Lynott et al. (2014) - Potential error in statistics reporting

### 2018-02-10 manuscript.Rmd
Adressed editor comment 5 * Also made minor revisions elsewhere in the same paragraph.

In order to distinguish between these possibilities, we can define what a “meaningful effect” (SESOI) would be and use it as the bounds for an equivalence test. In reality, we should do this before looking at the results of the study. If grades for this test were set at every `r printnum(sesoi * 100)`\% increase in correct answers (F = 0% to 6.25% ... A+ = 93.75% to 100%) we could decide that we are only interested in test score differences that correspond to an increase or decrease of at least 1 grade point. Thus, our SESOI becomes a difference in raw scores of `r printnum(sesoi * 100)`\%, or `r printnum(sesoi, digits = 4)`. We can then perform an equivalence test for a two-sample Welch’s *t*-test, with equivalence bounds of $\pm$ the SESOI of `r printnum(sesoi, digits = 4)`. The two one-sided-tests procedure consists of two one-sided tests, and we calculate *t*-values and *p*-values both for test against $\Delta_{L}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t1)`$, $p = `r printp(Moon$TOST_p1)`$, and $\Delta_{U}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t2)`$, $p = `r printp(Moon$TOST_p2)`$. While the *t*-test against $\Delta_{U}$ indicates that we can reject differences at least as extreme as `r sesoi`, the test against $\Delta_{L}$ shows we can *not* reject effects as large or larger than `r -sesoi`. The equivalence test is non-significant, which means we cannot reject a true effect as large or larger than 6.25% (Figure 2A), $t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`$, $p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))`$ (note we typically only report the one-sided test yielding the highest *p*-value in the result section).^[Because this is a replication of a study, it would have been reasonable to assume that we only want to reject effects in the same direction as the effect in the original study. After all, much smaller effects may indicate non-replication, but so do large effects in the opposite direction. In that case, we would perform an inferiority test (see figure 1D) against $\Delta_{U}$. As can be seen in Figure 2A, the 90\% CI does not overlap with $\Delta_{U}$, so we can reject effects as large or larger than $`r printnum(sesoi, digits = 4)`$.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions. Further studies are needed, which can be analyzed using a (small-scale) meta-analysis. The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both. Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L} = -0.1$ and $\Delta_{U} = 0.1$).

>>>>

In order to distinguish between these possibilities, we can define what a “meaningful effect” (SESOI) would be and use it as the bounds for an equivalence test. In reality, we should do this before looking at the results of the study. If grades for this test were set at every `r printnum(sesoi * 100)`\% increase in correct answers (F = 0% to 6.25% ... A+ = 93.75% to 100%) we could decide that we are only interested in test score differences that correspond to an increase or decrease of at least 1 grade point. Thus, our SESOI becomes a difference in raw scores of `r printnum(sesoi * 100)`\%, or `r printnum(sesoi, digits = 4)`. We can then perform an equivalence test for a two-sample Welch’s *t*-test, with equivalence bounds of $\pm$ the SESOI of `r printnum(sesoi, digits = 4)`. The two one-sided-tests procedure consists of two one-sided tests, and we calculate *t*-values and *p*-values both for test against $\Delta_{L}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t1)`$, $p = `r printp(Moon$TOST_p1)`$, and $\Delta_{U}$, $t(`r printnum(Moon$TOST_df)`) = `r printnum(Moon$TOST_t2)`$, $p = `r printp(Moon$TOST_p2)`$. While the *t*-test against $\Delta_{U}$ indicates that we can reject differences at least as extreme as `r printnum(sesoi, digits = 4)`, the test against $\Delta_{L}$ shows we can *not* reject effects as extreme or more extreme than `r printnum(-sesoi, digits = 4)`. The equivalence test is therefore non-significant, which means we cannot reject a true effect as large or larger than 6.25% (Figure 2A), $t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`$, $p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))`$ (note we typically only report the one-sided test yielding the highest *p*-value in the result section).^[Because this is a replication of a study, it would have been reasonable to assume that we only want to reject effects in the same direction as the effect in the original study. After all, much smaller effects may indicate non-replication, but so do large effects in the opposite direction. In that case, we would perform an inferiority test (see figure 1D) against $\Delta_{U}$. As can be seen in Figure 2A, the 90\% CI does not overlap with $\Delta_{U}$, so we can reject effects as large or larger than $`r printnum(sesoi, digits = 4)`$.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions. Further studies are needed, which can be analyzed using a (small-scale) meta-analysis. The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both. Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L} = -0.1$ and $\Delta_{U} = 0.1$).
